SIG Proceedings Paper in LaTeX Format

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/evaluation.tex
review

\subsection{Evaluating word embedings}
Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier for example. We cannot split our data set into train and test set. As the task that the network is learning is of no interest to us. Therefore, to ensure the quality of word embeddings other techniques have to be used.To define quality we first need to define a measure of similarity between two vectors. For this task we will use the cosine distance. Two vectors will be judged similiar if they point in the same direction.

\subsubsection{Word similarity and wordsim353}
Too measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
 We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature. The data set consists of 353 pairs of words rated by humans on their similarity. The similarity score is in the range of 1 and 10, with 10 being the best score. An example for two of such pairs can be found in Table \ref{fig:ws353_ex}. We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is the common procedure. This concludes our description of the evaluation methods of word embeddings. The next section will present the configuration of our network used to optimize the SGNS.
\iffalse
\begin{table}[tb]\centering
\caption{Example of pairs and their rating in wordsim353}
\begin{tabular}{l l l }
        \toprule
Word1 & Word2 & Score \\ \hline
        \midrule%

\textquote{FBI} & \textquote{Investigation} & 8.31 \\ \hline
\textquote{Mars} & \textquote{scientist} & 5.63 \\ \hline
        \midrule%
\end{tabular}
\label{fig:ws353_ex}
\end{table}
\fi

\subsection{Configuration of the network}\label{ssec:config}
The SGNS has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. The network was configured the following way: 
\begin{itemize}
\item \texttt{Negative Samples:} - \textit{10}
\item \texttt{Context Window:}  - \textit{5}
\item\texttt{Dimension of the embedding}: - \textit{100}
\item \texttt{Batch size}: - \textit{2000}
\item \texttt{Alpha}:  - \textit{(1e-5,1)}
\end{itemize}
This concludes our overview of the possible parameters of our SGNS, in the next section the reader will be given a description about the process of input shuffling.

@@

Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch. The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in order to find the best optimum. Furthermore, it's easier for the neural network to escape a local minimum. As for example if a network, without using shuffling,  had converged to a local minimum after one epoch it probably cannot escape it. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the SGNS. Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{sec:contribution}. When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples. But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller. In consequence the accuracy of the model should be optimized hence, leading to a better convergence time. A definition of convergence time is given in the next section.

To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset. To evaluate the resulting word embeddings we defined the cosine distance, which is 1 minuse the cosine of the angle of two vectors. We will measure the quality of the embeddings by analyzing the correlation of the cosine distance of pairs of vectors to humanly judged similarity score of the corresponding words, and used the wordsim353 to collect these scores. The reader was also given an explanation about the configuration of our network and introduced to the concept of input shuffling. With all this knowledge we can know present the achieved experimental results in the next section.

>>>>

\subsection{Evaluating word embedings}
Evaluating word embedding is not an easy task, such as evaluating the accuracy of a common classifier for example. We cannot split our data set into train and test set. As the task that the network is learning is of no interest to us. Therefore, to ensure the quality of word embeddings other techniques have to be used. To define quality we first need to define a measure of similarity between two vectors. For this task we will use the cosine distance. Two vectors will be judged similiar if they point in the same direction.

\subsubsection{Word similarity and wordsim353}
To measure the quality of our word embedding, in question of word similarity we must use a dataset to compare our self too.
 We chose wordsim353\footnote{http://www.cs.technion.ac.il/~gabr/resources/data/wordsim353/wordsim353.zip} for this task, as it's the most used in the related literature. The data set consists of 353 pairs of words rated by humans on their similarity. The similarity score is in the range of 1 and 10, with 10 being the best score. An example for two of such pairs can be found in Table \ref{fig:ws353_ex}. We will rank our embeddings on the Pearson correlation coefficient between the cosine distance and the scores attributed by humans, as this is the common procedure. This concludes our description of the evaluation methods of word embeddings. The next section will present the configuration of our network used to optimize the SGNS.
\begin{table}[tb]\centering
\caption{Example of pairs and their rating in wordsim353}
\begin{tabular}{l l l }
        \toprule
Word1 & Word2 & Score \\ \hline
        \midrule%

\textquote{FBI} & \textquote{Investigation} & 8.31 \\ \hline
\textquote{Mars} & \textquote{scientist} & 5.63 \\ \hline
        \midrule%
\end{tabular}
\label{fig:ws353_ex}
\end{table}

\subsection{Configuration of the network}\label{ssec:config}
The SGNS has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. The network was configured the following way: 
\begin{itemize}
\item \texttt{Negative Samples:} - \textit{10}
\item \texttt{Context Window:}  - \textit{5}
\item\texttt{Dimension of the embedding}: - \textit{100}
\item \texttt{Batch size}: - \textit{2000}
\item \texttt{Alpha}:  - \textit{(1e-5,1)}
\end{itemize}
This concludes our overview of the possible parameters of our SGNS, in the next subsection the reader will be given a description about the process of input shuffling.

@@

Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch. The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in order to find the best optimum. Furthermore, it's easier for the neural network to escape a local minimum. As for example if a network, without using shuffling,  had converged to a local minimum after one epoch it probably cannot escape it. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the SGNS. Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{sec:contribution}. When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples. But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller. In consequence the accuracy of the model should be optimized, hence leading to a better convergence time.

To summarize the evaluation, we deleted words that appear very often (subsampling) and very infrequently (min count) from our dataset, i.e text8 dataset. To evaluate the resulting word embeddings we defined the cosine distance, which is 1 minus the cosine of the angle of two vectors. We will measure the quality of the embeddings by analyzing the correlation of the cosine distance of pairs of vectors to humanly judged similarity score of the corresponding words, and used the wordsim353 to collect these scores. The reader was also given an explanation about the configuration of our network and introduced to the concept of input shuffling. With all this knowledge we can know present the achieved experimental results in the next section.

### 2019-09-28 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/related_work.tex
review

\section{Related Work}\label{sec:related_work}
In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm. This process does not favor parallelization. To deal with this specific problem Mikolov et al.\citep{mikolov2} used a Hogwild! approach proposed by Recht et al.\citep{hogwild}. The idea is to allow multiple threads to access a shared memory, in this case, the single model. In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by a one thread. The threads run parallel and have access to the shared memory. Therefore, overwriting errors are bound to happen. According to Recht et al.\citep{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough. But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words. There were several attempts at solving this issue, and optimize the throughput of the SGM.

@@

\subsection{Parallization in shared and Distributed Memory}
In contrast to the above approach, this one, introduced by Ji et al. \citep{intel} focused on the SGNS. The focus of this work was to speed up the computation of the loss function of the SGNS described in Equation \ref{eq:obj_neg_samples}. Where the dot product of the center word  with negative samples as well as the context word is computed. The dot product is a level 1-BLAS operation and is not as efficiently implemented into hardware, as matrix multiplication which are a level 3-BLAS operation.  To achieve such a transformation, Ji et al. decided to use the same negative samples for given amount of training samples. The training samples that will have the same negative samples are chosen as follows: given a word $w_c$ one can construct all the training samples where $w_c$ is a context word. The following training batch $X$ will result: $X = \{({w_{in}}^i, w_c) | w_c~appears~in~the~context~of~{w_{in}}^i\}$. The matrix multiplication can be represented in the following way:
\[
\begin{bmatrix}
w_c \\
w_{n_1} \\
\vdots \\
w_{n_k}\\
\end{bmatrix}
*
\begin{bmatrix}
w_{in}^1\\
\vdots\\
{w_{in}}^{2m}\\
\end{bmatrix}
\]
where, $w_{n_1}...w_{n_k}$ are the shared negative samples. The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch. Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\ % of accuracy. The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \citep{gpu} that parellelized the SGNS on GPU's.

\subsection{Accelleration of word2vec by using GPU's}
Seulki and Youngmin \citep{gpu} focused on getting a better throughput on the SGM with the use of GPU's. As the SGM is a sequential algorithm, it is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples. As the algorithm goes sequentially over a sentence, the samples next to each other, in order of execution, will almost every time have the same input word. Consequently, it's very hard to parallelize at this level. 
Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \citep{gpu} tried to parallelize their update. They achieved this by mapping each dimension to a CUDA thread while mapping each sentence to a CUDA block. As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem. If the execution time of the GPU kernel is greater than the time used to read the sentences, it could be a smart choice to use multiple GPU's. According to Seulki and Youngmin \citep{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance. They achieved their best results with 2 concurrent GPU'S. The acomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al. \citep{mikolov2}. The speedup was achieved, while maintaining the accuracy of the model. The problem with this and all the above optimization is that the code is not easily available or of use. Therefore, we need an optimized implementation of the SGM that is easily available. This is provided by Gensim \citep{gensim}, which will be outlined in the next subsection.
%%%%%%%%%%%
\subsection{Gensim}\label{ssec:gensim}
Gensim \citep{gensim} is a pure Python library that holds a state of the art implementation of the SGM. Gensim is written in Cython, which first allowed Gensim to have the same runtime as the original C code. Furthermore, it made use of BLAS's and precomputed sigmoid tables, while also trying to parallelize the training of different sentences. This finally yielded in a 4x speedup in run time compared to the orignal C code from Mikolov et al. Gensim as a python library, and because it was already used in related work \citep{intel} made the comparison easy and of value. \\ This concludes our overview of  related work. The next Section will cover our implementation of the Skip-Gram Model with negative sampling.

>>>>

\section{Related Work}\label{sec:related_work}
In the original model, the optimization is done with Stochastic Gradient Descent (SGD), which is a sequential algorithm. This process does not favor parallelization. To deal with this specific problem Mikolov et al.\citep{mikolov2} used a Hogwild! approach proposed by Recht et al.\citep{hogwild}. The idea is to allow multiple threads to access a shared memory, in this case, the single model. In the original SGM, the threads are constructed as follows: at the beginning of training, the dataset is split into $N$ evenly sized chunks and each of these chunks will be processed by one thread. The threads run parallel and have access to the shared memory. Therefore, overwriting errors are bound to happen. According to Recht et al.\citep{hogwild} the overwriting errors won't lead to a significant accuracy loss if the data is sparse enough. But in the case of NLP, the problem seems to be a bit more significant, and especially for word embedding, as many words share the same context words. There were several attempts at solving this issue, and optimize the throughput of the SGM.

@@

\subsection{Parallization in shared and Distributed Memory}
In contrast to the above approach, this one, introduced by Ji et al. \citep{intel} focused on the SGNS. The focus of this work was to speed up the computation of the loss function of the SGNS described in Equation \ref{eq:obj_neg_samples}. In this Equation the dot product of the center word  with negative samples as well as the context word is computed. The dot product is a level 1-BLAS operation and is not as efficiently implemented into hardware, as matrix multiplication which are a level 3-BLAS operation.  To achieve such a transformation, Ji et al. decided to use the same negative samples for given amount of training samples. The training samples that will have the same negative samples are chosen as follows: given a word $w_c$ one can construct all the training samples where $w_c$ is a context word. The following training batch $X$ will result: $X = \{({w_{in}}^i, w_c) | w_c~appears~in~the~context~of~{w_{in}}^i\}$. The matrix multiplication can be represented in the following way:
\[
\begin{bmatrix}
w_c \\
w_{n_1} \\
\vdots \\
w_{n_k}\\
\end{bmatrix}
*
\begin{bmatrix}
w_{in}^1\\
\vdots\\
{w_{in}}^{2m}\\
\end{bmatrix}
\]
where, $w_{n_1}...w_{n_k}$ are the shared negative samples. The above Matrix multiplication will compute the loss for each training sample and the weights of the used vectors will be updated after each batch. Ji et al.'s method achieves a 3.6 fold increase in throughput, by only losing 1\% of accuracy. The presented results were achieved on CPU's, as many modern machine learning libraries make use of the parallelizing capabilities of GPU's for matrix multiplications, the next subsection covers the work done by Seulki and Youngmin \citep{gpu} that parellelized the SGNS on GPU's.

\subsection{Accelleration of word2vec by using GPU's}
Seulki and Youngmin \citep{gpu} focused on getting a better throughput on the SGM with the use of GPU's. As the SGM is a sequential algorithm, it is not easy to parallelize it, especially if one wants to parallelize the training of individual training samples. As the algorithm goes sequentially over a sentence, the samples next to each other, in order of execution, will almost every time have the same input word. Consequently, it's very hard to parallelize at this level. 
Since the dimensions of the word embeddings are completly independent of each other  Seulki and Youngmin \citep{gpu} tried to parallelize their update. They achieved this by mapping each dimension of a weight vector represinting a word embedding(WE) to a CUDA thread while mapping each sentence to a CUDA block. As each CUDA block runs independently, the training of the sentences is parallelized, and the fact that sentences have different length is of no problem. If the execution time of the GPU kernel is greater than the time used to read the sentences, it could be a smart choice to use multiple GPU's. According to Seulki and Youngmin \citep{gpu}, if multiple GPU's are used, there is a need for synchronizing the model, which will hinder run time performance. They achieved their best results with 2 concurrent GPU'S. The acomplished results were very good as they managed a 20x speedup compared to a single threaded CPU execution, which is a 3x increase in comparison to the original C code from Mikolov et al. \citep{mikolov2}. The speedup was achieved, while maintaining the accuracy of the model. The problem with this and all the above optimization is that the code is not easily available or of use. Therefore, we need an optimized implementation of the SGM that is easily available. This is provided by Gensim \citep{gensim}, which will be outlined in the next subsection.
%%%%%%%%%%%
\subsection{Gensim}\label{ssec:gensim}
Gensim \citep{gensim} is a pure Python library that holds a state of the art implementation of the SGM. Gensim is written in Cython, which first allowed Gensim to have the same runtime as the original C code. Furthermore, it made use of BLAS's and precomputed sigmoid tables, while also trying to parallelize the training of different sentences. This finally yielded in a 4x speedup in run time compared to the orignal C code from Mikolov et al. Gensim as a python library, and because it was already used in related work \citep{intel} made the comparison easy and of value. \\ This concludes our overview of  related work. The next Section will cover our implementation of the Skip-Gram Model with negative sampling.
